---
title: "Lab2"
author: "Xintong Cai"
date: "2/4/2024"
output:
  pdf_document: default
  html_document: default
---

##  1.Descriptive statistics

```{r}
gactt <- readRDS('GACTT_RESULTS_ANONYMIZED_HW2.RDS')
table1 <- table(gactt$favorite_drink)
table1
```


```{r barplot}
barplot(table1)
```

We then try a pie chart.
```{r}
pie(table1)
```
Let's clean the data now. Remove the n/a in favorite drink and a_preference with scores equal to -9.
```{R}
df <- gactt[complete.cases(gactt$favorite_drink) & gactt$a_preference != -9 & gactt$preferred_strength != -9,] 
df$fav_is_pour <- ifelse(df$favorite_drink == 'Pourover','Pourover','Others')
```
Let's compare how the pourover compares with other favorite drinks since it accounts for the most proportion.
```{r}
table(df$fav_is_pour)
pie(table(df$fav_is_pour))
```

According to the summary of a_preference, we can see that the mean of a_preference is 3.33 and the max is 5. The mean of preferred_strength is 3.609.
```{r}
print(summary(df$a_preference))
print(summary(df$preferred_strength))
```

## Let's visualize the distribution of the strength of preference. 

```{r cars}
barplot(table(factor(df$preferred_strength)))
```

## 2. The second questioin: T-test

The fav_is_pour shows whether the favorite drink is pourover or other drinks.

```{r}
t_test_result <- t.test(a_preference ~ fav_is_pour, data = df)
t_test_result
```
According to the result, there is significant difference between the two groups.People who love pourover are more likely to enjoy the tasting lot. 

## 3. Third question: ANOVA 

(1) Let's set a new column. And 1/2 refer to light, 3 refers to medium, 4&5 refer to strong.
```{r}
df$preferred <- ifelse(df$preferred_strength %in% c(1,2),'light',
                       ifelse(df$preferred_strength %in% c(4,5),'strong','medium'))

```
(2) Run the anova test
```{r}
anova_result <- aov(a_preference ~ preferred_strength, data = df)
anova_result
```
According to the anova test, the estimated effects may be unbalanced, which shows that it might not be a perfect estimator. Moreover, the residual standard error is 1.18, which is pretty large considering the the scale is only 1-5. 


## 4.OLS Regression

### (1)
```{r}
lm_model <- lm(a_preference ~ fav_is_pour, data = df)
summary(lm_model)
```
According to the regression model, whether favorite is pourover has a big effect on the a_preference score. It agrees with the T-test we had in the second question saying that there's significant difference in those two groups: 'Pull over' and 'Others'.

### (2) The the numeric variable

```{r}
lm_numeric <- lm(a_preference ~ preferred_strength, data=df)
summary(lm_numeric)
```
According to the test, the preferred coffee strength doesn't have significant relationship with the a_preference. The result aligns with the anova test. 

It's hard to choose between ANOVA and regression because both can help in some way.
Anova is more suitable to study the difference between groups, considering the 1-5 as factors. Regression is more suitable to study the relationship between variables. The key is whether we should treat the preferred_strength as the continuous variable or discrete variable.

To choose between ANOVA test and the regression, I think in this case ANOVA might be a better choice after we have recoded the preferred_strength as three categorical variables. This may not be linear relationship between the two variables.



## Bonus task

```{r}
one_sample_t_test <- function(data, mu, alternative = "two.sided") {
  n <- length(data)
  x_bar <- mean(data)
  s <- sd(data)
  
  t_stat <- (x_bar - mu) / (s / sqrt(n))
  df <- n - 1
  
  if (alternative == "less") {
    p_value <- pt(t_stat, df)
  } else if (alternative == "greater") {
    p_value <- 1 - pt(t_stat, df)
  } else {
    p_value <- 2 * pt(abs(t_stat), df, lower.tail = FALSE)
  }
  result <- list(
    t_statistic = t_stat,
    degrees_of_freedom = df,
    p_value = p_value,
    alternative = alternative,
    method = "t-test"
  )
  
  return(result)
}

```

This is the function of the t-test. Let's try some examples.
We can choose to conduct one-tail or two-tail test by specifying the variable of alternative as 'less' or 'greater' or 'two.sided'.

```{r}
set.seed(200)
data <- rnorm(40, mean = 5, sd = 2)
mu <- 4.5
result <- one_sample_t_test(data, mu, alternative = "two.sided")
print(result)
```



